---
title: "Data Prep"
author: "Joey Stanley"
date: "February 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Previously, the data was prepped in the same script that combined all the spreadsheets. I think it might be good to have a stand-alone data prep file for Shiny, just so it doesn't get lost.

```{r}
library(tidyverse)

# remotes::install_github("joeystanley/joeyr", force = TRUE)
library(joeyr) # for tidy_mahalanobis and find_outliers
library(beepr)
```

# Where is all this data?

```{r}
# The DASS data
fave_path     <- "data/preprocessed/MFA_FAVE_210101_corrected_791_3.csv"

# The extra southern speakers
lags_path     <- "data/preprocessed/LAGS2020_MFA_FAVE_200904.csv"


# darla_path    <- "data/preprocessed/DASS_DARLA_191010.csv"

# Output path
shiny_path    <- "data/DASS_and_LAGS.csv"
```




# Read in and process the data

Instead of starting with the individual spreadsheets, I'll just read in the completed version.

```{r}
current_shiny_data <- read_csv("data/DASS_and_LAGS.csv", n_max = 10)
names(current_shiny_data)
```

Note that a lot of this processing could be done after FAVE and DARLA have been combined, but I felt it was easier to just to them separately and combine once they're all spic-and-span.

Here's how the filtering is defined:

* filter_stdev: TRUE = within 3 stdevs on F1 *and* F2. FALSE = outside 3 stdevs and should be removed
* filter_joey:  TRUE if kept via Joey's method. FALSE = removed via Joey's method and should be removed.
* filter_mahal: TRUE if within 95% Mahalanobis distance (Renwick's technique): FALSE = removed


Pretty sure this is the same as in the joeyr package, but it's here for reproducibility.

```{r}
find_outliers2 <- function(..., keep = 0.95) {
  
  # Capture that arbitrary list of variables.
  vars <- list(...)
  # Turn it into a dataframe
  df <- as.data.frame(vars, col.names = letters[1:length(vars)])
  # Figure out how many variables will be used.
  n_vars <- length(df)
  # Add a new column
  df$is_outlier <- FALSE
  df$id <- 1:nrow(df)
  
  # Get the length of the first (or any) of them
  total_n <- length(vars[[1]])
  # Figure out how many to remove
  n_to_remove <- floor(total_n * (1-keep))
  
  # Don't do anything for groups less than 20 tokens
  if (n_to_remove > 0) {
    
    # Loop through and remove the points one at a time.
    for (i in 1:n_to_remove) {
      
      # A new df with just the tokens not marked as outliers
      this_loop_df <- df[df$is_outlier == FALSE,]
      # A subset of that with just the columns that'll go into the mahalanobis() function
      data_for_mahal <- this_loop_df[,1:n_vars]
      
      # *Tips hat to Joe Fruehwald via Renwick for this line of code.*
      t_params <- MASS::cov.trob(data_for_mahal)
      this_loop_df$mahal_dist <- mahalanobis(data_for_mahal,
                                             center = t_params$center,
                                             cov    = t_params$cov)
      
      # # Add the mahalanobis distance to this iteration's dataframe. Sort so that the largest distance is first.
      this_loop_df <- this_loop_df[order(this_loop_df$mahal_dist, decreasing = TRUE),]
      
      # Extract the id of the largest distance.
      id_of_furthest_token <- this_loop_df$id[[1]]
      
      # Mark that one token as an outlier.
      df[df$id == id_of_furthest_token, "is_outlier"] <- TRUE
    }
    
  }
  
  # Return the vector that determines which points are outliers.
  df$is_outlier
}
```

## LAGS (10 speakers)

```{r}
lags_raw <- read_csv(lags_path,
                     col_types = cols(speaker = col_character(),
                                      kurath_type = col_character(),
                                      reel = col_character()))
```



```{r}
# Takes about 90 seconds because of find_outliers2()
start_time <- Sys.time()
lags <- lags_raw %>% 

  # Remove diphthongs
  filter(!vowel %in% c("AY", "AW", "OY", "ER")) %>%

  # Select the columns I want and rename them
  select(speaker, state, sex, ethnicity, birth_year,
         education_level, social_class, classification, kurath_type,
         vowel, stress, plt_vclass, word,
         plt_manner, plt_place, plt_voice, pre_seg, fol_seg,
         F1.50., F2.50., F1.50._lob, F2.50._lob,
         Bark_height, Bark_backness) %>%
  rename(ARPABET = vowel,
         Plotnik = plt_vclass, # plotnik codes, refers to vowel itself
         manner  = plt_manner, # manner of articulation of following consonant
         place   = plt_place,  # place of articulation of following consonant
         voice   = plt_voice   # voicing of following consonant
  ) %>%

  # Full names for the states
  mutate(state = fct_recode(state, "Georgia"="GA", "Tennessee"="TN",
                     "Alabama" = "AL", "Mississippi"="MS",
                     "Louisiana"="LA","Texas"="TX",
                     "Florida"="FL", "Arkansas"="AR"),
         stress = fct_recode(as.factor(stress),
                             "Primary"="1",
                             "Secondary"="2",
                             "No Stress"="0"),
         manner = fct_recode(manner, "rhotic"="central")) %>%

  # Filter for each vowel for each speaker
  group_by(speaker, Plotnik) %>%
  transform(x_norm = scale(F1.50.),
            y_norm = scale(F2.50.)) %>%
  mutate(filter_stdev = abs(x_norm) < 3 & abs(y_norm) < 3) %>%
  select(-x_norm, -y_norm) %>%
  as_tibble() %>%

  # Find outliers with my filter
  group_by(speaker, Plotnik) %>%
  filter(!is.na(F1.50.), !is.na(F1.50.)) %>%
  mutate(filter_joey = find_outliers2(F1.50., F2.50.),
         filter_mahal = tidy_mahalanobis(F1.50., F2.50.)) %>%
  ungroup() %>%
  
  print()
Sys.time() - start_time
beep()
```

## FAVE (DASS)

```{r}
fave_raw <- read_csv(fave_path,
                     col_types = cols(kurath_type = col_character(),
                                      reel = col_character()))
```

```{r}
fave_raw %>%
  count(speaker)
```




```{r}
# Takes about 6 minutes because of find_outliers2()
start_time <- Sys.time()
fave <- fave_raw %>% 

  # Remove diphthongs
  filter(!vowel %in% c("AY", "AW", "OY", "ER")) %>%

  # Select the columns I want and rename them
  select(speaker, state, sex, ethnicity, birth_year,
         education_level, social_class, classification, kurath_type,
         vowel, stress, plt_vclass, word,
         plt_manner, plt_place, plt_voice, pre_seg, fol_seg,
         F1.50., F2.50., F1.50._lob, F2.50._lob,
         Bark_height, Bark_backness) %>%
  rename(ARPABET = vowel,
         Plotnik = plt_vclass, # plotnik codes, refers to vowel itself
         manner  = plt_manner, # manner of articulation of following consonant
         place   = plt_place,  # place of articulation of following consonant
         voice   = plt_voice   # voicing of following consonant
  ) %>%
  
  # Full names for the states
  mutate(state = fct_recode(state, "Georgia"="GA", "Tennessee"="TN",
                     "Alabama" = "AL", "Mississippi"="MS",
                     "Louisiana"="LA","Texas"="TX",
                     "Florida"="FL", "Arkansas"="AR"),
         stress = fct_recode(as.factor(stress), 
                             "Primary"="1", 
                             "Secondary"="2", 
                             "No Stress"="0"),
         manner = fct_recode(manner, "rhotic"="central")) %>%

  # Filter for each vowel for each speaker
  group_by(speaker, Plotnik) %>%
  transform(x_norm = scale(F1.50.),
            y_norm = scale(F2.50.)) %>%
  mutate(filter_stdev = abs(x_norm) < 3 & abs(y_norm) < 3) %>%
  select(-x_norm, -y_norm) %>%
  as_tibble() %>%

  # Find outliers with my filter
  group_by(speaker, Plotnik) %>%
  filter(!is.na(F1.50.), !is.na(F1.50.)) %>%
  mutate(filter_joey = find_outliers2(F1.50., F2.50.),
         filter_mahal = tidy_mahalanobis(F1.50., F2.50.)) %>%
  ungroup() %>%
  
  print()
Sys.time() - start_time
beep()
```



## DARLA (not needed)

```{r, eval = FALSE}
darla_raw <- read_csv(darla_path, 
                      col_types = cols(kurath_type = col_character(),
                                       reel = col_character()))
```

```{r, eval = FALSE}
# Takes about 12 minutes because of clean_joey()
start_time <- Sys.time()
darla <- darla_raw %>% 

  # Remove diphthongs
  filter(!vowel %in% c("AY", "AW", "OY", "ER")) %>%

  # Select the columns I want and rename them
  select(speaker, state, sex, ethnicity, birth_year,
         education_level, social_class, classification, kurath_type,
         vowel, stress, plt_vclass, word,
         plt_manner, plt_place, plt_voice, pre_seg, fol_seg,
         F1.50., F2.50., F1.50._lob, F2.50._lob,
         Bark_height, Bark_backness) %>%
  rename(ARPABET = vowel,
         Plotnik = plt_vclass, # plotnik codes, refers to vowel itself
         manner  = plt_manner, # manner of articulation of following consonant
         place   = plt_place,  # place of articulation of following consonant
         voice   = plt_voice   # voicing of following consonant
  ) %>%
  
  # Full names for the states
  mutate(state = fct_recode(state, "Georgia"="GA", "Tennessee"="TN",
                     "Alabama" = "AL", "Mississippi"="MS",
                     "Louisiana"="LA","Texas"="TX",
                     "Florida"="FL", "Arkansas"="AR"),
         stress = fct_recode(as.factor(stress), 
                             "Primary"="1", 
                             "Secondary"="2", 
                             "No Stress"="0"),
         manner = fct_recode(manner, "rhotic"="central")) %>%

  # Filter for each vowel for each speaker
  group_by(speaker, Plotnik) %>%
  transform(x_norm = scale(F1.50.),
            y_norm = scale(F2.50.)) %>%
  mutate(filter_stdev = abs(x_norm) < 3 & abs(y_norm) < 3) %>%
  select(-x_norm, -y_norm) %>%
  as_tibble() %>%

  # Do my own filter as well
  group_by(speaker, Plotnik) %>%
  do(clean_joey(., probs = 0.95)) %>%
  ungroup() %>%
  
  print()
Sys.time() - start_time
```





# Join them together

Check to see if all the columns match.

```{r}
reduce(list(tibble(names = names(current_shiny_data),
                   old = TRUE),
            tibble(names = names(fave),
                   fave = TRUE),
            tibble(names = names(lags),
                   lags = TRUE)),
       full_join,
       by = "names")
```


```{r}
all <- bind_rows(list("LAGS" = lags,
                      "DASS" = fave),
                 .id = "corpus")
```

Double check again to make sure the corpus thing worked.

```{r}
reduce(list(tibble(names = names(current_shiny_data),
                   old = TRUE),
            tibble(names = names(fave),
                   dass = TRUE),
            tibble(names = names(lags),
                   lags = TRUE),
            tibble(names = names(all),
                   all = TRUE)),
       full_join,
       by = "names")
```


## Audit/Query/Quality Check

How much data is there?

```{r}
nrow(all[all$corpus=="DASS",])
nrow(all[all$corpus=="LAGS",])
```

```{r}
all %>%
  count(corpus, speaker) %>%
  pivot_wider(names_from = corpus, values_from = n)
```


# Export

```{r}
all %>% write_csv(shiny_path)
```



# Prepare maps data

```{r}
latlong <- read_csv("C:\\Users\\Atlas\\Google Drive\\NSF\\LAGS_latlong.csv") %>%
  mutate(speaker = str_pad(speaker, side = "left", pad = "0", width = 3)) %>%
  print()

# Get the number of vowels per speaker.
n_obs_per_speaker <- shinyData %>%
  group_by(speaker) %>%
  summarize(vowels = n()) %>%
  print()

shiny_metadata <- metadata %>%
  # Account for the presumed typos in the sector names.
  mutate(sector_name = factor(sector_name),
         sector_name = fct_recode(sector_name,
                                  "WF" = "GA",
                                  "EL"= "GM")) %>%
  select(-sector, -age_level) %>%
  left_join(latlong, by="speaker") %>%
  left_join(n_obs_per_speaker) %>%
  print()


# Write it out
write_csv(shiny_metadata, "C:\\Users\\Atlas\\Google Drive\\NSF\\ShinyApp\\data\\metadata.csv")
```
